Grip strength, mating success, and immune and energy costs in a population sample of US women and men: a registered report

### 2024-08-01 paper.qmd
Saved time consuming simulation results to a file. Edited figure captions to reflect combination of pilot and confirmatory results.

```{r}
source('functions.R')
source('pilot_analysis.R')
source('simulate.R')
source('sexspecificmodels.R')
source('stage2_analysis.R')
source('compare.R')
```

@@

```{r}
#| label: fig-ecdf
#| fig-width: 10
#| fig-cap: Cumulative distribution of the number of sex partners, by sex. Participants who reported 100 or more partners were removed from the analysis. The x-axis is on a log scale, so 1 was added to the number of partners to prevent removal of individuals with 0 sex partners (only for this plot and not for any analyses).

@@

```{r}
#| label: fig-coefs
#| fig-width: 10
#| fig-height: 8
#| fig-cap: "Coefficients of sex-specific strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables. Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions. Bars are 95\% CIs. Left: Pilot study results. Right: Confirmatory study results. For the full regression tables, see the SI."

plot_coefs + plot_coefs_stage2 + plot_layout(axes = 'collect_y', axis_titles = 'collect')
```

@@

```{r}
#| label: fig-allcoefs
#| fig-cap: Coefficients from all models of our four mating success outcomes and our five sets of potential confounding variables. Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions. Bars are 95\% CIs. For the full regression tables, see the SI.

plot_allcoefs
```

@@

plot_immune_both <- 
  plot_immune + guides(colour = 'none', shape = 'none') + plot_immune_stage2 + plot_layout(axes = 'collect', guides = 'collect')

ggsave("Figures/plot_immune_both.pdf", plot_immune, width = 14, height = 7)

plot_immune_both
```

@@

```{r}
#| label: fig-energy
#| fig.width: 12
#| fig.height: 8
#| fig.cap: Coefficient plot of predictors of dietary energy intake (kcals) and dietary protein (g) from generalized linear models (Pilot). Variables labelled (S) have been centered at the mean and standardized by 2 SD. Bars are 95% CIs.

plot_energy <- forestplot(
  intake_models$m_energy,
  intake_models$m_energy_alt,
  intercept = F,
  facet = F,
  dodgewidth = .4,
  varnames = vnames, 
  modelnames = mmnames) +
  scale_color_binary() +
  ggtitle('Stage 1: Pilot Energy') +
  theme_bw(17) +
  guides(colour = guide_none(), shape = guide_none())

plot_protein <- forestplot(
  intake_models$m_protein,
  intake_models$m_protein_alt,
  intercept = F,
  facet = F,
  dodgewidth = .4,
  varnames = vnames, 
  modelnames = mmnames) +
  scale_color_binary() +
  guides(colour = guide_legend(reverse = T), shape = guide_legend(reverse = T)) +
  ggtitle('Stage 1: Pilot Protein') +
  theme_bw(17) +
  theme(axis.text.y = element_blank())

p <- plot_energy + plot_protein + plot_layout(axes = 'collect') 
ggsave("Figures/plot_energy_protein.pdf", p, width = 12, height = 6)
p
```

@@

| Outcome            | Strength coefficient | Strength X Sex coefficient |
|--------------------|:--------------------:|:--------------------------:|
| Partnered status   |           +          |              -             |
| Lifetime partners  |           +          |              -             |
| Past year partners |           0          |              0             |
| Age at first sex   |           0          |              0             |
| Immunity           |           -          |              0             |
| Energy             |           0          |                            |
| Protein            |           +          |                            |

: Predictions. A plus sign (+) indicates a predicted significantly positive coefficient in all models, a negative sign (-) indicates a predicted significantly negative coefficient in all models, and zero (0) indicates a predicted non-significant coefficient in any model. A blank indicates not relevant. {#tbl-predictions}

@@

Models and R code for three of our four mating success outcome variables remain identical to those used in the Stage 1 pilot study. However, although the Poisson regression model of lifetime sexual partners in the pilot study included age as a control, it is more appropriate to instead include it as an exposure variable (an offset) in the model, which we now do. We also treat this exposure as years since sexual maturity (age - 12 years). The change in the strength and strength:sex coefficients from this change in model specification is minimal (see @sfig-comparison).

@@

```{r}
#| label: fig-stage2coef
#| fig-width: 10
#| fig-height: 8
#| fig-cap: Coefficients of sex-specific strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables. Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions. Bars are 95\% CIs. For the full regression tables, see the SI.

plot_coefs_stage2
```

:::{#fig-stage2effects}
```{r}
plot_effects_all
```

Effects of sex-specific standardized strength on mating outcomes, by sex. Strength is centered and scaled by 2 SD, separately for each sex (almost all participants therefore fall between -1 and 1). Plots are based on the anthropometric control models, with control variables held at their mean values. Lifetime partners is modeled as the rate of new partners per year since reaching sexual maturity (age 12).
:::

@@

Descriptive statistics.

::: {#stbl-stats}
`{r} tbl_stack(list(con, cat))`

Descriptive statistics of pilot study sample.
:::

@@

::: {#sfig-tradeoff}
```{r}
#| fig-width: 11
plot_lifetime_coefs
```

@@

### Effects plots (Pilot)

:::{#sfig-effects}
```{r}
effects_plots(models)
```

Effects of sex-specific standardized strength on mating outcomes, by sex. Strength is centered and scaled by 2 SD, separately for each sex (almost all participants therefore fall between -1 and 1). Plots are based on the anthropometric control models, with control variables held at their mean values. Lifetime partners is modeled as the rate of new partners per year since reaching sexual maturity (age 12).
:::

@@

cor_bothsexes <- ggcorrplot(cor_mat_cors, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Both sexes')
ggsave("Figures/cor_bothsexes.pdf", cor_bothsexes, width = 12, height = 12)
cor_bothsexes
```

@@

cor_females <- ggcorrplot(cor_mat_cors_f, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Females')
ggsave("Figures/cor_females.pdf", cor_females, width = 12, height = 12)
cor_females
```

@@

cor_males <- ggcorrplot(cor_mat_cors_m, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Males')
ggsave("Figures/cor_males.pdf", cor_males, width = 12, height = 12)
cor_males
```

@@

```{r}
#| results: asis
for (i in 1:nrow(modelsummaries)){
  cat(paste('####', modelsummaries$title[i]))
  print(modelsummaries$summary[[i]])
}
```

@@

```{r}
#| results: asis
for (i in 1:nrow(dfmodelsummaries)){
  cat(paste('####', dfmodelsummaries$title[i]))
  print(dfmodelsummaries$summary[[i]])
}
```

@@

@@

A comparison of strength and strength:sex cofficients with the original model specification for the lifetime sex partners (anthropometric controls) that included age as a control variable (orignal), and the specification that now instead includes years since sexual maturity (age-12) as an exposure variable (offset) (new). Both models are quasi-Poisson models.
:::

>>>>

```{r}
source('functions.R')
source('pilot_analysis.R')
source('sexspecificmodels.R')
source('stage2_analysis.R')
source('compare.R')

# source('simulate.R')
load("df_pwr.rda")
```

@@

```{r}
#| label: fig-ecdf
#| fig-width: 10
#| fig-cap: Cumulative distribution of the number of sex partners, by sex. Participants who reported 100 or more partners were removed from the analysis. The x-axis is on a log scale, so 1 was added to the number of partners to prevent removal of individuals with 0 sex partners (only for this Figure and not for any analyses).

@@

```{r}
#| label: fig-coefs
#| fig-width: 10
#| fig-height: 8
#| fig-cap: "Coefficients of sex-specific strength, and strength X sex interaction for our four mating success outcomes and our five sets of potential confounding variables. Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions. Bars are 95% CIs. For the full regression tables, see the SI."

# plot_coefs + plot_coefs_stage2 + plot_layout(axes = 'collect_y', axis_titles = 'collect')
plot_coefs_combined
```

@@

```{r}
#| label: fig-allcoefs
#| fig-cap: Stage 2 Confirmatory study coefficients from all models of the four mating success outcomes and the five sets of potential confounding variables. Partnered coefficients are from logistic regressions, partner number coefficients are from quasi-Poisson glm regressions, and age at first sex coefficients are from Gaussian glm regressions. Mexican Amercians are the base level for the ethnicity control variable. Bars are 95\% CIs. For the full regression tables, see the SI.

plot_allcoefs_stage2
```

@@

# plot_immune_both <- 
#   plot_immune + guides(colour = 'none', shape = 'none') + plot_immune_stage2 + plot_layout(axes = 'collect', guides = 'collect')

ggsave("Figures/plot_immune_strength_combined.pdf", plot_immune_strength_combined, width = 14, height = 7)

plot_immune_strength_combined
```

@@

```{r}
#| label: fig-energy
#| fig.width: 12
#| fig.height: 8
#| fig.cap: Coefficient plots of predictors of dietary energy (kcals) and protein (g) intake from generalized linear models for Stage 1 Pilot (left) and Stage 2 Confirmatory (right). Variables labelled (S) have been centered at the mean and standardized by 2 SD. Bars are 95% CIs.

# p <- plot_energy_coefs / plot_protein_coefs + plot_layout(guides = 'collect', axis_titles = 'collect')
ggsave("Figures/plot_intake_strength_combined.pdf", plot_intake_strength_combined, width = 12, height = 6)
plot_intake_strength_combined
```

@@

| Outcome            | Strength coefficient  | Strength X Sex coefficient |
|--------------------|:---------------------:|:--------------------------:|
| Partnered status   |           +   ✓      |              -   ✓         |
| Lifetime partners  |           +   ✓      |              -   ❌         |
| Past year partners |           0   ✓      |              0   ✓         |
| Age at first sex   |           0   ✓      |              0   ✓         |
| Immunity           |           -   ❌      |              0   ✓         |
| Energy             |           0   ❌      |                            |
| Protein            |           +   ✓      |                            |

: Predictions. A plus sign (+) indicates a predicted significantly positive coefficient in all models, a negative sign (-) indicates a predicted significantly negative coefficient in all models, and zero (0) indicates a predicted non-significant coefficient in any model. A blank indicates not relevant. ✓: Supported in confirmatory study. ❌: Not supported in confirmatory study. {#tbl-predictions}

@@

Models and R code for three of our four mating success outcome variables remain identical to those used in the Stage 1 pilot study. However, although the Poisson regression model of lifetime sexual partners in the pilot study included age as a control, it is more appropriate to instead include age as an exposure variable (an offset) in the model, which we now do. We also treat this exposure as years since sexual maturity (age - 12 years). The changes in the strength and strength:sex coefficients from this change in model specification are minimal (see @sfig-comparison).

@@

```{r}
#| label: fig-effects
#| fig-width: 10
#| fig-height: 8
#| fig-cap: Effects of sex-specific standardized strength on mating outcomes, by sex. Strength is centered and scaled by 2 SD, separately for each sex (almost all participants therefore fall between -1 and 1). Plots are based on the anthropometric control models, with control variables held at their mean values. Lifetime partners is modeled as the rate of new partners per year since reaching sexual maturity (age 12).

plot_effects_all
```

@@

::: {#stbl-stats}
```{r}
pilot_summary_stats <- summary_tables(designsG$d.design.adults)
tbl_stack(list(pilot_summary_stats$con, pilot_summary_stats$cat))
```

Pilot: Descriptive statistics and sex differences for participants ages 18-60 using population weights
:::

@@

::: {#sfig-tradeoff}
```{r}
#| fig-width: 11
plot_lifetime_coefs_combined
```

@@

@@

pilot_cor_mats <- cor_mats(designsG)

cor_bothsexes <- ggcorrplot(pilot_cor_mats$cor_mat_cors, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Pilot: Both sexes')
ggsave("Figures/cor_bothsexes.pdf", cor_bothsexes, width = 12, height = 12)
cor_bothsexes
```

@@

cor_females <- ggcorrplot(pilot_cor_mats$cor_mat_cors_f, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Pilot: Females')
ggsave("Figures/cor_females.pdf", cor_females, width = 12, height = 12)
cor_females
```

@@

cor_males <- ggcorrplot(pilot_cor_mats$cor_mat_cors_m, lab = T, hc.order = T, hc.method = 'ward.D', title = 'Pilot: Males')
ggsave("Figures/cor_males.pdf", cor_males, width = 12, height = 12)
cor_males
```

@@

```{r}
#| results: asis
for (i in 1:nrow(modelsummaries)){
  cat(paste('####', str_replace(modelsummaries$title[i], '\n', ' '))) # Remove line feed char
  print(modelsummaries$summary[[i]])
}
```

@@

```{r}
#| results: asis
for (i in 1:nrow(dfmodelsummaries)){
  cat(paste('####', str_replace(dfmodelsummaries$title[i], '\n', ' ')))
  print(dfmodelsummaries$summary[[i]])
}
```

@@

::: {#stbl-confirmstats}
```{r}
confirm_summary_stats <- summary_tables(designsH2$d.design.adults)
tbl_stack(list(confirm_summary_stats$con, confirm_summary_stats$cat))
```

Confirmatory: Descriptive statistics and sex differences for participants ages 18-60 using population weights
:::

@@

A comparison of strength and strength:sex cofficients with the original model specification for the lifetime sex partners (anthropometric controls) that included age as a control variable (original), and the specification that now instead includes years since sexual maturity (age-12) as an exposure variable (offset) (new). Both models are quasi-Poisson models.
:::
